---
title: "2022 Regionals Predictive Modeling"
author: "Emily Murphy, Dwight Wynne"
date: "6/14/2022"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r setup2, include = FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(lattice)
library(caret)
library(tidymodels)
library(parsnip)
library(dplyr)
library(gt)
Softball2015 <- read.csv("Data/Softball2015.csv")

Softball2016 <- read.csv("Data/Softball2016.csv")

Softball2017 <- read.csv("Data/Softball2017.csv")

Softball2018 <- read.csv("Data/Softball2018.csv")


#this is the google spreadsheet
Regionals <- read.csv("Data/NCAA Tournament Games 2015-2018 - Sheet1 new.csv")
#filters each data set into teams that went to Regionals
NCAA2015 <- filter(Softball2015, NCAA %in% c("At-large","Auto"))
NCAA2016 <- filter(Softball2016, NCAA %in% c("At-large","Auto"))
NCAA2017 <- filter(Softball2017, NCAA %in% c("At-large","Auto"))
NCAA2018 <- filter(Softball2018, NCAA %in% c("At-large","Auto"))

#adds year variable
NCAA2015 <- mutate(NCAA2015, Year=2015)
NCAA2016 <- mutate(NCAA2016, Year=2016)
NCAA2017 <- mutate(NCAA2017, Year=2017)
NCAA2018 <- mutate(NCAA2018, Year=2018)

#cannot row bind unless number of columns are equal; this removes extra columns in 2018 data set
NCAA2018 <- subset(NCAA2018, select=-c(PitcherK, PitcherBB))

#row binds the four datasets
RegionalStats <- data.frame(rbind(NCAA2015,NCAA2016,NCAA2017,NCAA2018))

#reorders columns so year is first
RegionalStats <- RegionalStats[, c(32,1:31)]

#Adds stats: fielding percentage, singles, total bases, double plays per game, slugging percentage, success rate of stolen bases
#second line adds batting average
RegionalStats <- mutate(RegionalStats, FieldingPct = (PO+A)/(PO+A+E), Singles = H - Doubles - Triples - HR, 
                        TB = Singles + 2*Doubles + 3*Triples + 4*HR, DPPerGame = DP/G, 
                        SlgPct = TB/AB, SuccessRate = SB/(SB+CS))
RegionalStats <- mutate(RegionalStats, BA=H/AB)

#Joins Regionals and stats into one data frame by game
RegionalGames <- Regionals %>% left_join(RegionalStats, by = c("Home.Team" = "Name", "Year"),  
                                         suffix = c(".x", ".home")) %>% left_join(RegionalStats, 
                                                                                  by = c("Visiting.Team" = "Name", "Year"), suffix = c(".home",".visit"))

#Adds column indicating home team win
RegionalGames <- mutate(RegionalGames, Home.Win = ifelse(Home.Score>Visiting.Score, "Yes", "No"))

#Adds column indicating run differental (absolute value of Home-Visiting runs)
RegionalGames <- mutate(RegionalGames, Run.Diff = Home.Score-Visiting.Score)

#Creates new data frame that standardizes important variables
RegionalGames_Std <- mutate(RegionalGames, Run.Diff = Home.Score-Visiting.Score,
                            Singles.home = Singles.home/G.home, Singles.visit=Singles.visit/G.visit,
                            TB.home=TB.home/G.home, TB.visit=TB.visit/G.visit, 
                            Doubles.home=Doubles.home/G.home, Doubles.visit=Doubles.visit/G.visit,
                            Triples.home=Triples.home/G.home, Triples.visit=Triples.visit/G.visit,
                            HR.home=HR.home/G.home, HR.visit=HR.visit/G.visit,
                            SB.home=SB.home/G.home, SB.visit=SB.visit/G.visit,
                            CS.home=CS.home/G.home, CS.visit=CS.visit/G.visit,
                            RunsScored.home=RunsScored.home/G.home, RunsScored.visit=RunsScored.visit/G.visit,
                            PO.home=PO.home/G.home, PO.visit=PO.visit/G.visit,
                            A.home=A.home/G.home, A.visit=A.visit/G.visit,
                            E.home=E.home/G.home, E.visit=E.visit/G.visit,
                            RunsAllowed.home=RunsAllowed.home/G.home, RunsAllowed.visit=RunsAllowed.visit/G.visit)

#Adds column indicating if home team or visiting team (or neither) is regional host
RegionalGames_Std <- mutate(RegionalGames_Std, 
                            Host= case_when(Regional.Host==Home.Team~1, 
                                            Regional.Host==Visiting.Team~-1,
                                            Regional.Host != Home.Team & Regional.Host != Visiting.Team ~ 0) )
Differences <- mutate(RegionalGames_Std, Doubles = Doubles.home - Doubles.visit, 
                      Triples = Triples.home - Triples.visit,
                      HR = HR.home - HR.visit, 
                      RunsScored = RunsScored.home - RunsScored.visit,
                      SB = SB.home - SB.visit, 
                      CS = CS.home - CS.visit, 
                      IP = IP.home - IP.visit, 
                      RunsAllowed = RunsAllowed.home - RunsAllowed.visit,
                      ER = ER.home - ER.visit,
                      ERA = ERA.home - ERA.visit,
                      PO = PO.home - PO.visit,
                      A = A.home - A.visit,
                      E = E.home - E.visit,
                      DP = DPPerGame.home - DPPerGame.visit,
                      FieldingPct = FieldingPct.home - FieldingPct.visit, 
                      Singles = Singles.home - Singles.visit,
                      TB = TB.home - TB.visit, 
                      SlgPct = SlgPct.home - SlgPct.visit,
                      SuccessRate = SuccessRate.home - SuccessRate.visit,
                      BA = BA.home - BA.visit,
                      ConferenceDiff = case_when(ConfNumBids.home > 1 & ConfNumBids.visit == 1 ~ 1, ConfNumBids.home == 1 & ConfNumBids.visit > 1 ~ -1, TRUE ~ 0 ))


Differences$Home.Win <- as.factor(Differences$Home.Win)
Softball2019 <- read.csv("Data/Softball2019.csv")

Softball2021 <- read.csv("Data/Softball2021.csv")

Softball2022 <- read.csv("Data/Softball2022.csv")



#read in spreadsheets for 2019, 2021, 2022
Games2019 <- read.csv("Data/NCAA 2019 Regional Games.csv")
Games2021 <- read.csv("Data/NCAA 2021 Regional Games.csv")
Games2022 <- read.csv("Data/NCAA 2022 Regional Games.csv")

#filters each data set into teams that went to Regionals
NCAA2019 <- filter(Softball2019, NCAA %in% c("At-large","Auto"))
NCAA2021 <- filter(Softball2021, NCAA %in% c("Yes"))
NCAA2022 <- filter(Softball2022, NCAA %in% c("Yes"))

#gets rid of extra columns in 2022
#NCAA2022 <- subset(NCAA2022, select=-c(X, X.1, X.2, X.3, X.4, X.5, X.6, X.7, X.8, X.9, X.10, X.11, X.12, X.13, X.14, X.15, X.16, X.17, X.18))

#adds year variable
NCAA2019 <- mutate(NCAA2019, Year=2019)
NCAA2021 <- mutate(NCAA2021, Year=2021)
NCAA2022 <- mutate(NCAA2022, Year=2022)

#moves year column to front
NCAA2019 <- NCAA2019[, c(35,1:34)]
NCAA2021 <- NCAA2021[, c(36,1:35)]
NCAA2022 <- NCAA2022[, c(36,1:35)]

#adds stats to match RegionalGames_Std
NCAA2019 <- mutate(NCAA2019, FieldingPct = (PO+A)/(PO+A+E), Singles = H - Doubles - Triples - HR, 
                        TB = Singles + 2*Doubles + 3*Triples + 4*HR, DPPerGame = DP/G, 
                        SlgPct = TB/AB, SuccessRate = SB/(SB+CS), BA=H/AB)
NCAA2021 <- mutate(NCAA2021, FieldingPct = (PO+A)/(PO+A+E), Singles = H - Doubles - Triples - HR, 
                   TB = Singles + 2*Doubles + 3*Triples + 4*HR, DPPerGame = DP/G, 
                   SlgPct = TB/AB, SuccessRate = SB/(SB+CS), BA=H/AB)
NCAA2022 <- mutate(NCAA2022, FieldingPct = (PO+A)/(PO+A+E), Singles = H - Doubles - Triples - HR, 
                   TB = Singles + 2*Doubles + 3*Triples + 4*HR, DPPerGame = DP/G, 
                   SlgPct = TB/AB, SuccessRate = SB/(SB+CS), BA=H/AB)

#joins game data with stat data by team name for both home and visiting
Regionals2019 <- Games2019 %>% left_join(NCAA2019, by = c("Home.Team" = "Name", "Year"),  
                                                          suffix = c(".x", ".home")) %>% left_join(NCAA2019, 
                                                                                                   by = c("Visiting.Team" = "Name", "Year"), suffix = c(".home",".visit"))

Regionals2021 <- Games2021 %>% left_join(NCAA2021, by = c("Home.Team" = "Name", "Year"),  
                                          suffix = c(".x", ".home")) %>% left_join(NCAA2021, 
                                                                                   by = c("Visiting.Team" = "Name", "Year"), suffix = c(".home",".visit"))

Regionals2022 <- Games2022 %>% left_join(NCAA2022, by = c("Home.Team" = "Team", "Year"),  
                                          suffix = c(".x", ".home")) %>% left_join(NCAA2022, 
                                                                                   by = c("Visiting.Team" = "Team", "Year"), suffix = c(".home",".visit"))

#Adds column indicating yes/no for home team win
Regionals2019 <- mutate(Regionals2019, Home.Win = ifelse(Home.Score>Visiting.Score, "Yes", "No"))

Regionals2021 <- mutate(Regionals2021, Home.Win = ifelse(Home.Score>Visiting.Score, "Yes", "No"))

Regionals2022 <- mutate(Regionals2022, Home.Win = ifelse(Home.Score>Visiting.Score, "Yes", "No"))

#Creates new data frame that standardizes important variables
Regionals2019 <- mutate(Regionals2019, Run.Diff = Home.Score-Visiting.Score,
                            Singles.home = Singles.home/G.home, Singles.visit=Singles.visit/G.visit,
                            TB.home=TB.home/G.home, TB.visit=TB.visit/G.visit, 
                            Doubles.home=Doubles.home/G.home, Doubles.visit=Doubles.visit/G.visit,
                            Triples.home=Triples.home/G.home, Triples.visit=Triples.visit/G.visit,
                            HR.home=HR.home/G.home, HR.visit=HR.visit/G.visit,
                            SB.home=SB.home/G.home, SB.visit=SB.visit/G.visit,
                            CS.home=CS.home/G.home, CS.visit=CS.visit/G.visit,
                            RunsScored.home=RunsScored.home/G.home, RunsScored.visit=RunsScored.visit/G.visit,
                            PO.home=PO.home/G.home, PO.visit=PO.visit/G.visit,
                            A.home=A.home/G.home, A.visit=A.visit/G.visit,
                            E.home=E.home/G.home, E.visit=E.visit/G.visit,
                            RunsAllowed.home=RunsAllowed.home/G.home, RunsAllowed.visit=RunsAllowed.visit/G.visit)

Regionals2021 <- mutate(Regionals2021, Run.Diff = Home.Score-Visiting.Score,
                        Singles.home = Singles.home/G.home, Singles.visit=Singles.visit/G.visit,
                        TB.home=TB.home/G.home, TB.visit=TB.visit/G.visit, 
                        Doubles.home=Doubles.home/G.home, Doubles.visit=Doubles.visit/G.visit,
                        Triples.home=Triples.home/G.home, Triples.visit=Triples.visit/G.visit,
                        HR.home=HR.home/G.home, HR.visit=HR.visit/G.visit,
                        SB.home=SB.home/G.home, SB.visit=SB.visit/G.visit,
                        CS.home=CS.home/G.home, CS.visit=CS.visit/G.visit,
                        RunsScored.home=RunsScored.home/G.home, RunsScored.visit=RunsScored.visit/G.visit,
                        PO.home=PO.home/G.home, PO.visit=PO.visit/G.visit,
                        A.home=A.home/G.home, A.visit=A.visit/G.visit,
                        E.home=E.home/G.home, E.visit=E.visit/G.visit,
                        RunsAllowed.home=RunsAllowed.home/G.home, RunsAllowed.visit=RunsAllowed.visit/G.visit)

Regionals2022 <- mutate(Regionals2022, Run.Diff = Home.Score-Visiting.Score,
                        Singles.home = Singles.home/G.home, Singles.visit=Singles.visit/G.visit,
                        TB.home=TB.home/G.home, TB.visit=TB.visit/G.visit, 
                        Doubles.home=Doubles.home/G.home, Doubles.visit=Doubles.visit/G.visit,
                        Triples.home=Triples.home/G.home, Triples.visit=Triples.visit/G.visit,
                        HR.home=HR.home/G.home, HR.visit=HR.visit/G.visit,
                        SB.home=SB.home/G.home, SB.visit=SB.visit/G.visit,
                        CS.home=CS.home/G.home, CS.visit=CS.visit/G.visit,
                        RunsScored.home=RunsScored.home/G.home, RunsScored.visit=RunsScored.visit/G.visit,
                        PO.home=PO.home/G.home, PO.visit=PO.visit/G.visit,
                        A.home=A.home/G.home, A.visit=A.visit/G.visit,
                        E.home=E.home/G.home, E.visit=E.visit/G.visit,
                        RunsAllowed.home=RunsAllowed.home/G.home, RunsAllowed.visit=RunsAllowed.visit/G.visit)

#adds host variable
Regionals2019 <- mutate(Regionals2019, 
                            Host= case_when(Regional.Host==Home.Team~1, 
                                            Regional.Host==Visiting.Team~-1,
                                            Regional.Host != Home.Team & Regional.Host != Visiting.Team ~ 0) )

Regionals2021 <- mutate(Regionals2021, 
                        Host= case_when(Regional.Host==Home.Team~1, 
                                        Regional.Host==Visiting.Team~-1,
                                        Regional.Host != Home.Team & Regional.Host != Visiting.Team ~ 0) )

Regionals2022 <- mutate(Regionals2022, 
                        Host= case_when(Regional.Host==Home.Team~1, 
                                        Regional.Host==Visiting.Team~-1,
                                        Regional.Host != Home.Team & Regional.Host != Visiting.Team ~ 0) )

#creates new dataframe that includes the difference between important stats
Differences2019 <- mutate(Regionals2019, Doubles = Doubles.home - Doubles.visit, 
                      Triples = Triples.home - Triples.visit,
                      HR = HR.home - HR.visit, 
                      RunsScored = RunsScored.home - RunsScored.visit,
                      SB = SB.home - SB.visit, 
                      CS = CS.home - CS.visit, 
                      IP = IP.home - IP.visit, 
                      RunsAllowed = RunsAllowed.home - RunsAllowed.visit,
                      ER = ER.home - ER.visit,
                      ERA = ERA.home - ERA.visit,
                      PO = PO.home - PO.visit,
                      A = A.home - A.visit,
                      E = E.home - E.visit,
                      DP = DPPerGame.home - DPPerGame.visit,
                      FieldingPct = FieldingPct.home - FieldingPct.visit, 
                      Singles = Singles.home - Singles.visit,
                      TB = TB.home - TB.visit, 
                      SlgPct = SlgPct.home - SlgPct.visit,
                      SuccessRate = SuccessRate.home - SuccessRate.visit,
                      BA = BA.home - BA.visit)

Differences2021 <- mutate(Regionals2021, Doubles = Doubles.home - Doubles.visit, 
                          Triples = Triples.home - Triples.visit,
                          HR = HR.home - HR.visit, 
                          RunsScored = RunsScored.home - RunsScored.visit,
                          SB = SB.home - SB.visit, 
                          CS = CS.home - CS.visit, 
                          IP = IP.home - IP.visit, 
                          RunsAllowed = RunsAllowed.home - RunsAllowed.visit,
                          ER = ER.home - ER.visit,
                          ERA = ERA.home - ERA.visit,
                          PO = PO.home - PO.visit,
                          A = A.home - A.visit,
                          E = E.home - E.visit,
                          DP = DPPerGame.home - DPPerGame.visit,
                          FieldingPct = FieldingPct.home - FieldingPct.visit, 
                          Singles = Singles.home - Singles.visit,
                          TB = TB.home - TB.visit, 
                          SlgPct = SlgPct.home - SlgPct.visit,
                          SuccessRate = SuccessRate.home - SuccessRate.visit,
                          BA = BA.home - BA.visit)

Differences2022 <- mutate(Regionals2022, Doubles = Doubles.home - Doubles.visit, 
                          Triples = Triples.home - Triples.visit,
                          HR = HR.home - HR.visit, 
                          RunsScored = RunsScored.home - RunsScored.visit,
                          SB = SB.home - SB.visit, 
                          CS = CS.home - CS.visit, 
                          IP = IP.home - IP.visit, 
                          RunsAllowed = RunsAllowed.home - RunsAllowed.visit,
                          ER = ER.home - ER.visit,
                          ERA = ERA.home - ERA.visit,
                          PO = PO.home - PO.visit,
                          A = A.home - A.visit,
                          E = E.home - E.visit,
                          DP = DPPerGame.home - DPPerGame.visit,
                          FieldingPct = FieldingPct.home - FieldingPct.visit, 
                          Singles = Singles.home - Singles.visit,
                          TB = TB.home - TB.visit, 
                          SlgPct = SlgPct.home - SlgPct.visit,
                          SuccessRate = SuccessRate.home - SuccessRate.visit,
                          BA = BA.home - BA.visit)


#final model
test6 <- glm(Home.Win ~ Host+HR +
                                RunsScored +
                                RunsAllowed +
                                DP +
                                SlgPct +
                                FieldingPct +
                                Singles +
                                Doubles +
                                Triples +
                                SuccessRate +
                                BA + ConferenceDiff, family = binomial, data = Differences) %>% MASS::stepAIC()

Differences2019 <- mutate(Differences2019, ConferenceDiff = case_when(ConfNumBids.home > 1 & ConfNumBids.visit == 1 ~ 1, ConfNumBids.home == 1 & ConfNumBids.visit > 1 ~ -1, TRUE ~ 0 ))
Differences2021 <- mutate(Differences2021, ConferenceDiff = case_when(ConfNumBids.home > 1 & ConfNumBids.visit == 1 ~ 1, ConfNumBids.home == 1 & ConfNumBids.visit > 1 ~ -1, TRUE ~ 0 ))
Differences2022 <- mutate(Differences2022, ConferenceDiff = case_when(ConfNumBids.home > 1 & ConfNumBids.visit == 1 ~ 1, ConfNumBids.home == 1 & ConfNumBids.visit > 1 ~ -1, TRUE ~ 0 ))

Statistic <- c("CSUF 2022 Actual", "Estimated 2022 Goals", "Changing RA in Model")
Chances <- c(0.1151719, 0.2503823, 0.203444)
compare2022AZ <- data.frame(Statistic, Chances)
compareAZ2 <- compare2022AZ[c(1,2),]


Chances_LSU <- c(0.2477817, 0.4580774, 0.3925945)
compare2022LSU <- data.frame(Statistic, Chances_LSU)


Chances_SD <- c(0.4403116, 0.6687409, 0.6068669)
compare2022SD <- data.frame(Statistic, Chances_SD)

```

The statistics from the 2015-2018 seasons were merged with the data from the 2015-2018 Regional games to create one set of data that provides each Regional game for a particular year, who the host, home, and visiting teams were, who won each game, and each team's statistics for that season.

```{r echo = FALSE}
RegionalsTable <- RegionalGames_Std[c(1,163,229,359), ] %>% rename(`Regional Host` = Regional.Host, `Game Number` = Game.Number, `Home Team` = Home.Team, `Visiting Team` = Visiting.Team, `Home Score` = Home.Score, `Visiting Score` = Visiting.Score, `Home Team Win?` = Home.Win)
RegionalsTable %>%
  gt() %>%
  cols_align(
    align = "center"
  ) %>%
  tab_header(
    title = md("**Regional Games and Team Statistics** Years 2015 - 2018"),
    subtitle = "Figure 1: A sample of rows from the dataset. Home team stats are labeled .home and visiting team stats are labeled .visit"
  ) %>%
  opt_align_table_header(align = "left") %>%
  tab_options(table.border.top.width=3, table.border.top.color="#00274C", table.border.bottom.width=3,table.border.bottom.color="#00274C", 
              column_labels.border.bottom.color = "#00274C", table_body.hlines.color="#DC8218", column_labels.font.weight="bold")
```


All statistics were used and a model was built that kept 8 different stats that were shown to be most important when predicting who wins a game. The model also takes the regional host into account and adjusts for conference.

The model resulted in a 75% accuracy when tested on 2022 games, meaning it accurately predicted the outcome of approximately 75% of the 2022 Regional games. Now, using this model, the chances of a home team winning a game given both the home and visiting team statistics can be predicted with about 75% confidence.

Given CSUF's 2022 statistics, the model can be used to estimate the chances CSUF had of winning each game in the Arizona St. regional. This can be compared to the chances CSUF would have had according to the model if they had reached what the estimated goals would have been for each statistic this season. These estimated goals were calculated by averaging the 2021 WCWS per game statistics, and then averaging those with CSUF's 2021 per game statistics to obtain a goal for each stat. Again, each statistic was changed to be per game rather than the total to make for easier comparison. 

Figure 1 below shows CSUF's actual 2022 data in the 8 statistics that were considered to have the most influence on who wins a game, as well as what the estimated goals were for this season in the same stats.

```{r, echo = FALSE}

statistic <- c("CSUF Actual", "Estimated goals", "Changing runs allowed")
BA <- c(0.31, 0.32, 0.31)
RunsAllowed <- c(3.95, 2.63, 2.95)
Singles <- c(6.14, 5.30, 6.14)
Doubles <- c(1.39, 1.46, 1.39)
Triples <- c(0.09, 0.14, 0.09)
HR <- c(0.70, 1.30, 0.70)
SlgPct <- c(0.44, 0.54, 0.44)
DP <- c(0.29, 0.27, 0.29)

goals <- data.frame(statistic, BA, RunsAllowed, Singles, Doubles, Triples, HR, SlgPct, DP)

goalstable <- goals %>% rename(`Batting Average` = BA, `Runs Allowed` = RunsAllowed, 
                               `Home Runs` = HR, `Slugging Percentage` = SlgPct, 
                               `Double Plays` = DP,`Input` = statistic)
goalstable1 <- goalstable[c(1,2),]
goalstable1 %>% 
  gt(caption = "Figure 1: A table showing the statistics the model determined were most effective in predicting which team wins. \nIncludes CSUF's actual season values in these statistics as well as the estimated goals for each.  ") %>%
  cols_align(
    align = "center") %>%
  tab_header(
    title = md("Predictive Model Statistics"), 
    subtitle = "2022 Season"
  ) %>%
  tab_options(table.border.top.width=3, table.border.top.color="#00274C", table.border.bottom.width=3,table.border.bottom.color="#00274C", 
              column_labels.border.bottom.color = "#00274C", table_body.hlines.color="#DC8218", column_labels.font.weight="bold")
```


The following graphs compare CSUF's chances of winning according to the model using CSUF's 2022 stats and what their stats would have been if they had achieved the estimated goals.
```{r est-goals, echo = FALSE }
par(mfrow=c(1,3))
SD <- ggplot(data = compare2022SD[compare2022SD$Statistic!= "Changing RA in Model",], aes(x = Statistic, y = Chances_SD)) + 
  geom_bar(stat = "identity", width = 0.5, fill = "#FF7900") +
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0))+
  labs(title = "Odds of Winning vs. San Diego St.", x = "  ", y = "Odds of CSUF Winning", caption = "Figure 2.1: Note that the chanc eof winning increases by ~23%")+
  scale_y_continuous(labels = percent, breaks = seq(0,1.0, by = 0.25))+
  coord_cartesian(ylim = c(0,1.0)) +
  geom_text(aes(label = percent(Chances_SD, .11)), 
            vjust=-0.3, size = 3.5, color = "#00274C")
SD

LSU <- ggplot(data = compare2022LSU[compare2022LSU$Statistic!= "Changing RA in Model",], aes(x = Statistic, y = Chances_LSU)) + 
  geom_bar(stat = "identity", width = 0.5, fill = "#FF7900") +
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0))+
  labs(title = "Odds of Winning vs. LSU", x = "  ", y = "Odds of CSUF Winning", caption="Figure 2.2: Note that the chance of winning increases by ~21%")+
  scale_y_continuous(labels = percent, breaks = seq(0,1.0, by = 0.25))+
  coord_cartesian(ylim = c(0,1.0)) +
  geom_text(aes(label = percent(Chances_LSU, .11)), 
            vjust=-0.3, size = 3.5, color = "#00274C")
LSU

Arizona.St <- ggplot(data = compare2022AZ[compare2022AZ$Statistic!= "Changing RA in Model",], aes(x = Statistic, y = Chances)) + 
  geom_bar(stat = "identity", width = 0.5, fill = "#FF7900") +
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0))+
  labs(title = "Odds of Winning vs. Arizona St.", x = "  ", y = "Odds of CSUF Winning", caption="Figure 2.3: Note that the chance of winning increases by ~13.5%, but remains well under 50%")+
  scale_y_continuous(labels = percent, breaks = seq(0,1.0, by = 0.25))+
  coord_cartesian(ylim = c(0,1.0)) +
  geom_text(aes(label = percent(Chances, .11)), 
            vjust=-0.3, size = 3.5, color = "#00274C")
Arizona.St
```

According to the model, CSUF's statistics for the 2022 season gave them approximately a 44% and 25% chance of beating San Diego St. and LSU respectively. If CSUF had achieved the estimated goals for the season, the model predicts that their chances of winning against the same teams would have increased to 67% and 46%.

Against Arizona St., however, achieving the estimated goals for the season only would have resulted in about a 25% chance of winning.

\newpage The model can be used to estimate the effect of improvement in one statistic on the overall chances of winning against a given team. For example, CSUF allowed about 4 runs per game. If pitching is something the team is planning on working on, then we can use the model to estimate the chances CSUF would have of winning against teams with similar stats to those CSUF faced in the regional by changing the number of runs they allow per game from 4 to 3 and keeping every other statistic the same. 

Figure 3 below is the same table as figure 1, but includes a row that outlines the change in runs allowed as well as how the other stats would be maintained.

```{r, echo = FALSE}
goalstable %>% 
  gt(caption = "Figure 3: Similar table to figure 1, but includes a third row \nshowing the data input into the model when changing runs allowed from 3.95 to 2.95") %>%
  cols_align(
    align = "center") %>%
  tab_style(
    style = list(
      cell_fill(color = "#00274C"),
      cell_text(weight = "bold", color = "white")
      ),
    locations = cells_body(
      columns = 3,
      rows = 3
    )) %>%
  tab_header(
    title = md("Example of Predictive Modeling Changing Runs Allowed")
  ) %>%
  tab_options(table.border.top.width=3, table.border.top.color="#00274C", table.border.bottom.width=3,table.border.bottom.color="#00274C", 
              column_labels.border.bottom.color = "#00274C", table_body.hlines.color="#DC8218", column_labels.font.weight="bold")
```


Figures 4.1-4.3 below are the same graphs as figures 2.1-2.3, but add a third column that indicates the change in CSUF's chances of winning against these three teams when changing only the number of runs they allow per game, leaving every other statistic equal to their 2022 season stats.

```{r model-goals, echo = FALSE}
actual_2022 <- Regionals2022[19:24,]

Differences_act <- mutate(actual_2022, Doubles = Doubles.home - Doubles.visit, 
                          Triples = Triples.home - Triples.visit,
                          HR = HR.home - HR.visit, 
                          RunsScored = RunsScored.home - RunsScored.visit,
                          SB = SB.home - SB.visit, 
                          CS = CS.home - CS.visit, 
                          IP = IP.home - IP.visit, 
                          RunsAllowed = RunsAllowed.home - RunsAllowed.visit,
                          ER = ER.home - ER.visit,
                          ERA = ERA.home - ERA.visit,
                          PO = PO.home - PO.visit,
                          A = A.home - A.visit,
                          E = E.home - E.visit,
                          DP = DPPerGame.home - DPPerGame.visit,
                          FieldingPct = FieldingPct.home - FieldingPct.visit, 
                          Singles = Singles.home - Singles.visit,
                          TB = TB.home - TB.visit, 
                          SlgPct = SlgPct.home - SlgPct.visit,
                          SuccessRate = SuccessRate.home - SuccessRate.visit,
                          BA = BA.home - BA.visit)
Differences_act <-  mutate(Differences_act, ConferenceDiff = case_when(ConfNumBids.home > 1 & ConfNumBids.visit == 1 ~ 1, ConfNumBids.home == 1 & ConfNumBids.visit > 1 ~ -1, TRUE ~ 0 ))



new_2022 <- Regionals2022[19:24,]
new_2022$RunsAllowed.visit[which(new_2022$Visiting.Team == "Cal St. Fullerton")] <- 2.946429
new_2022$RunsAllowed.home[which(new_2022$Home.Team == "Cal St. Fullerton")] <- 2.946429

Differences_new <- mutate(new_2022, Doubles = Doubles.home - Doubles.visit, 
                          Triples = Triples.home - Triples.visit,
                          HR = HR.home - HR.visit, 
                          RunsScored = RunsScored.home - RunsScored.visit,
                          SB = SB.home - SB.visit, 
                          CS = CS.home - CS.visit, 
                          IP = IP.home - IP.visit, 
                          RunsAllowed = RunsAllowed.home - RunsAllowed.visit,
                          ER = ER.home - ER.visit,
                          ERA = ERA.home - ERA.visit,
                          PO = PO.home - PO.visit,
                          A = A.home - A.visit,
                          E = E.home - E.visit,
                          DP = DPPerGame.home - DPPerGame.visit,
                          FieldingPct = FieldingPct.home - FieldingPct.visit, 
                          Singles = Singles.home - Singles.visit,
                          TB = TB.home - TB.visit, 
                          SlgPct = SlgPct.home - SlgPct.visit,
                          SuccessRate = SuccessRate.home - SuccessRate.visit,
                          BA = BA.home - BA.visit)
Differences_new <-  mutate(Differences_new, ConferenceDiff = case_when(ConfNumBids.home > 1 & ConfNumBids.visit == 1 ~ 1, ConfNumBids.home == 1 & ConfNumBids.visit > 1 ~ -1, TRUE ~ 0 ))

est_2022 <- Regionals2022[19:24,]

#replacing CSUF visit data with estimated goals
est_2022$HR.visit[which(est_2022$Visiting.Team == "Cal St. Fullerton")] <- 1.3039820
est_2022$RunsAllowed.visit[which(est_2022$Visiting.Team == "Cal St. Fullerton")] <- 2.631781
est_2022$DPPerGame.visit[which(est_2022$Visiting.Team == "Cal St. Fullerton")] <- 0.2656308
est_2022$SlgPct.visit[which(est_2022$Visiting.Team == "Cal St. Fullerton")] <- 0.5390819
est_2022$Singles.visit[which(est_2022$Visiting.Team == "Cal St. Fullerton")] <- 5.301730
est_2022$Doubles.visit[which(est_2022$Visiting.Team == "Cal St. Fullerton")] <- 1.456112
est_2022$Triples.visit[which(est_2022$Visiting.Team == "Cal St. Fullerton")] <- 0.1414057
est_2022$BA.visit[which(est_2022$Visiting.Team == "Cal St. Fullerton")] <- 0.3195273

#replacing CSUF home data with estimated goals
est_2022$HR.home[which(est_2022$Home.Team == "Cal St. Fullerton")] <- 1.3039820
est_2022$RunsAllowed.home[which(est_2022$Home.Team == "Cal St. Fullerton")] <- 2.631781
est_2022$DPPerGame.home[which(est_2022$Home.Team == "Cal St. Fullerton")] <- 0.2656308
est_2022$SlgPct.home[which(est_2022$Home.Team == "Cal St. Fullerton")] <- 0.5390819
est_2022$Singles.home[which(est_2022$Home.Team == "Cal St. Fullerton")] <- 5.301730
est_2022$Doubles.home[which(est_2022$Home.Team == "Cal St. Fullerton")] <- 1.456112
est_2022$Triples.home[which(est_2022$Home.Team == "Cal St. Fullerton")] <- 0.1414057
est_2022$BA.home[which(est_2022$Home.Team == "Cal St. Fullerton")] <- 0.3195273

#recreating Differences data on regional to apply model
Differences_est <- mutate(est_2022, Doubles = Doubles.home - Doubles.visit, 
                                             Triples = Triples.home - Triples.visit,
                                             HR = HR.home - HR.visit, 
                                             RunsScored = RunsScored.home - RunsScored.visit,
                                             SB = SB.home - SB.visit, 
                                             CS = CS.home - CS.visit, 
                                             IP = IP.home - IP.visit, 
                                             RunsAllowed = RunsAllowed.home - RunsAllowed.visit,
                                             ER = ER.home - ER.visit,
                                             ERA = ERA.home - ERA.visit,
                                             PO = PO.home - PO.visit,
                                             A = A.home - A.visit,
                                             E = E.home - E.visit,
                                             DP = DPPerGame.home - DPPerGame.visit,
                                             FieldingPct = FieldingPct.home - FieldingPct.visit, 
                                             Singles = Singles.home - Singles.visit,
                                             TB = TB.home - TB.visit, 
                                             SlgPct = SlgPct.home - SlgPct.visit,
                                             SuccessRate = SuccessRate.home - SuccessRate.visit,
                                             BA = BA.home - BA.visit)
Differences_est <-  mutate(Differences_est, ConferenceDiff = case_when(ConfNumBids.home > 1 & ConfNumBids.visit == 1 ~ 1, ConfNumBids.home == 1 & ConfNumBids.visit > 1 ~ -1, TRUE ~ 0 ))

csuf_actual <- test6 %>% predict(Differences_act, type = "response")
csuf_est <- test6 %>% predict(Differences_est, type = "response")
csuf_new <- test6 %>% predict(Differences_new, type = "response")
Statistic <- c("CSUF 2022 Actual", "Estimated 2022 Goals", "Changing RA in Model")
Chances <- c(0.1151719, 0.2503823, 0.1142600)
compare2022AZ <- data.frame(Statistic, Chances)
compareAZ2 <- compare2022AZ[c(1,2),]

Chances_LSU <- c(0.2477817, 0.4580774, 0.3177409)
compare2022LSU <- data.frame(Statistic, Chances_LSU)

Chances_SD <- c(0.4403116, 0.6687409, 0.4734231)
compare2022SD <- data.frame(Statistic, Chances_SD)

```

```{r, echo = FALSE}

SD2 <- ggplot(data = compare2022SD, aes(x = Statistic, y = Chances_SD, fill = Statistic)) + 
  geom_bar(stat = "identity", width = 0.75) +
  scale_fill_manual(values = c("#FF7900", "#00274C", "#00274C"))+
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5), legend.position = "none")+
  labs(title = "Odds of Winning vs. San Diego St. Using Model", x = "  ", y = "Odds of CSUF Winning", caption = "Figure 4.1")+
  scale_y_continuous(labels = percent, breaks = seq(0,1.0, by = 0.25))+
  coord_cartesian(ylim = c(0,1.0))+
  geom_text(aes(label = percent(Chances_SD, .11)), 
            vjust=-0.3, size = 3.5, color = "#00274C")+
  scale_x_discrete(limits = c("CSUF 2022 Actual", "Estimated 2022 Goals", "Changing RA in Model"))
SD2

LSU2 <- ggplot(data = compare2022LSU, aes(x = Statistic, y = Chances_LSU, fill = Statistic)) + 
  geom_bar(stat = "identity", width = 0.75) +
  scale_fill_manual(values = c("#FF7900", "#00274C", "#00274C"))+
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5), legend.position = "none")+
  labs(title = "Odds of Winning vs. LSU Using Model", x = "  ", y = "Odds of CSUF Winning", caption = "Figure 3.2")+
  scale_y_continuous(labels = percent, breaks = seq(0,1.0, by = 0.25))+
  coord_cartesian(ylim = c(0,1.0))+
  geom_text(aes(label = percent(Chances_LSU, .11)), 
            vjust=-0.3, size = 3.5, color = "#00274C")+
  scale_x_discrete(limits = c("CSUF 2022 Actual", "Estimated 2022 Goals", "Changing RA in Model"))
LSU2

Arizona.St2 <- ggplot(data = compare2022AZ, aes(x = Statistic, y = Chances, fill = Statistic)) + 
  geom_bar(stat = "identity", width = 0.75) +
  scale_fill_manual(values = c("#FF7900", "#00274C", "#00274C"))+
  theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5), legend.position = "none")+
  labs(title = "Odds of Winning vs. Arizona St. Using Model", x = "  ", y = "Odds of CSUF Winning", caption = "Figure 4.3")+
  scale_y_continuous(labels = percent, breaks = seq(0,1.0, by = 0.25))+
  coord_cartesian(ylim = c(0,1.0))+
  geom_text(aes(label = percent(Chances, .11)), 
            vjust=-0.3, size = 3.5, color = "#00274C")+
  scale_x_discrete(limits = c("CSUF 2022 Actual", "Estimated 2022 Goals", "Changing RA in Model"))
Arizona.St2

```

By changing the number of runs CSUF allows by 1 less run, we can see how that would effect the likelihood of CSUF winning each of these games, keeping the other statistics constant. Using the model in this way can provide immediate feedback on goals for next season and gives the ability to set goals that are attainable for the team.
